\section{Introduction}

Since the beginning of recorded time we have wondered whether we are
alone or whether there are other planets that might support life. The
scientific quest for extrasolar or exoplanets (planets beyond our own
solar system) began in the mid 19th century, but it was only as recent
as 1992 that astronomers have been able to confirm the existence of
other planetary systems.  Since then {\tt describe resources,
  telescopes, etc} have been devoted to identifying stars with
orbiting planets, and in particular, those that might have earth-like
planets. As of June 2010 over 450 exoplanets have been detected, the
majority using techniques that rely on measurements of a star's radial
velocity.  The presence of a planet orbiting a star will lead to a
periodic ``wobble'' in the the star's radial velocity, while if there
are no planets present the expected velocity is constant although
measurements fluctuate because of stellar jitter and other sources of
noise.  Based on a short sequence of noisy measurements, the challenge
is to determine whether there are no planets or one, two or more
planets, and if there are planets present to characterize their orbital
parameters.  {\tt cite work of Gregory and Ford, but mention
  limitations?}  If the data do not provide strong evidence either in
favor or against the presence of planets, additional observations may
be needed to provide confirmation.  Determining which stars to observe
and when to schedule scarce telescope time optimally is critical given
limited resources; Bayesian methods in particular are ideally suited
for updating information/evidence and decision
making in such sequential sampling problems.


The primary question of whether there are no planets or one or more
may be poised as a Bayesian model selection problem, where the
collection of models under consideration are the zero-planet model
$\M_0$, one-planet $\M_1$, two-planet $\M_2$, etc.  In Bayesian
statistics, the posterior probability of a model $\M_j$ requires
accurately calculating the integrated or marginal likelihood (see
\citet{Clyd:Geor:2004} for a general overview or
\citet{Clyd:etal:2005} for some of the issues that arise in
astronomy).  In the exoplanet application, each model $\M_p$ has
$2+5p$ parameters where $p$ is the number of planets, so even the
simplest single planet model requires a seven dimensional integral
over the parameter space to obtain the marginal likelihood.  Because
of the nonlinear nature of the models, analytic solutions do not
exist.  While low in dimension compared to many modern problems of
model selection, the models are highly nonlinear making Laplace
approximations untenable given the modest sample sizes and multi-modal
posterior distribution; constructing good proposal distributions for
Reversible Jump-Markov Chain Monte Carlo (RJ-MCMC) algorithms is near
impossible; just designing an efficient MCMC algorithm for sampling
from posterior distributions for a model with one planet is
challenging on its own due to the multi-modal nature of the posterior
distribution.  


Out of available methods for numerical integration, importance
sampling (IS) generally produces the best estimates of marginal
likelihoods in terms of bias and variance \citep{Christian's
  comparisons?}. The key to building an efficient IS algorithm is to
design an importance function that mimics the integrand.  Building
such a function can be quite challenging even in lower to modest
dimensional settings; with the multi-modal posterior distributions
that arise in the exoplanet setting the task is all the more
ambitious!  In this paper, we propose a novel adaptive method for
constructing an importance function using a mixture of Student-$t$
distributions that gradually evolves to the target posterior
distribution.  The paper is organized as follows.  We begin by
presenting an overview of the radial velocity models for exoplanet
detection in Section \ref{sec:exo} and discuss the scientific issues
that inspired this work. In Section \ref{sec:IS} we review importance
sampling in the context of calculating marginal likelihoods for
Bayesian model selection. In Section \ref{sec:AAIS} we present the
Adaptive Annealed Importance Sampling method developed for the
exoplanet problem. This method integrates several key features of
other methods: mixture models, annealing, and adaptation via a
sequence of distributions to build a flexible importance distribution
that mimics the posterior distribution, but incorporates ideas
successfully used in trans-dimensional methods to adapt the number of
kernels in the mixture by utilizing split and merge moves.  In Section
\ref{sec:simulation} we present two simulation studies that
demonstrate the efficiency of the method compared to other approaches.
The first involves a challenging flared helix in three dimensions,
while the second seven dimensional problem that was selected to reflex
the target function in the exoplanet problem.  In Section
\ref{sec:app} we illustrate the method in the analysis of several real
star systems with one and two planets. Finally, we conclude with
discussion of possible extensions of the method in Section
\ref{sec:disc}.
